### 2020-06-27 2020/WCGO_MCO_OptimizationLetters/2020_WCGO_MCO.tex
some minor corrections

\institute{               
Victor Gergel \at
%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{gergel@unn.ru} 
\and
Evgeniy Kozinov \at
%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{evgeny.kozinov@itmm.unn.ru} 
\and
Konstantin Barkalov \at
%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{konstantin.barkalov@itmm.unn.ru} \\
\at Institute of Informational Technologies, Mathematics and Mechanics, Mathematical Center, Lobachevsky State University of Nizhni Novgorod, Gagarin ave., 23, Nizhni Novgorod, Russia 
}

@@

For many years, global optimization has been the subject of many scientific studies. The results of research have been published in a large number of monographs -- see, for example, \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}. Making a reasonable choice of the approach for solving GO problems from the entire set of already existing global search algorithms requires taking into account the following:

\begin{itemize}
	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function at the point of the global minimum should be compared with the values of the function at all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain,
	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e. it must be dense only in the vicinity of globally optimal points,
	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- on one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}),
	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods. Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23},
	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
\end{itemize}

Given the requirements listed above, the proposed approach for solving MMCOlex problems is based on three key ideas:

\begin{itemize}
	\item Reduction of multidimensional multiextremal optimization to problems of one-dimensional global search by the space-filling curves;
	\item Using the global optimization methods that generate nonuniform coverages of the search domain;
	\item Intensive use of the search information obtained in the process of computations for solving multiple global optimization problems.
\end{itemize}

@@

The presence of nonlinear constraints significantly increasesthe complexity of global optimization problems; the resulting solutions must belong to a feasible domain
\begin{equation}\label{eq:17}
\Delta_g  = \{ x\in \Delta=[0,1]:g_i(x)\leq 0, 1 \leq i \leq m \}.
\end{equation}

@@

\begin{theorem}
If the AGCS algorithm is used to solve the problem (\ref{eq:16}), and the following conditions are satisfied:
\begin{enumerate}
	\item the functions $g_j(x)$, $1 \leq j \leq m+1$, $x\in[0,1]$, satisfy the H\"older condition (\ref{eq:13}) with the constants $H_j$, $1 \leq j \leq m+1$,
	\item for the values $H^k_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$, are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
\begin{equation} \label{eq:29}
r_\nu H^k_\nu > 2H_\nu, r_\nu > 1, 1 \leq \nu \leq m+1,
\end{equation}
\end{enumerate}
then the set of limit points of the sequence $\{x^k\}$ generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:16}) with $\varepsilon=0$ in the stopping criterion (\ref{eq:28}).
\end{theorem}

@@

\begin{theorem}
Let the next solved optimization problem $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previously solved problem $\phi(x)$ by some limited value $\delta$, i.e.
\begin{equation}\label{eq:32}
|\psi(x)-\phi(x)| \leq \delta, x \in [0,1]
\end{equation}
and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
\begin{equation}\label{eq:33}
|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
\end{equation}
if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
\begin{equation}\label{eq:34}
m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
\end{equation}
where
\begin{equation}\label{eq:35}
m=\max{r_l H_l^k}, 1 \leq l \leq M^k, k>k_0,
\end{equation}
\begin{equation}\label{eq:36}
L=\max{L_i}, 1 \leq i \leq s, L_i\; \text{from}\; (\ref{eq:3}),
\end{equation}
\begin{equation}\label{eq:37}
\psi(x^*)=\min{\psi(x)},x \in [0,1],
\end{equation}
\begin{equation}\label{eq:38}
x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k, 
\end{equation}
\end{theorem}(the points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively).

@@

\begin{theorem}
Let the optimization problem $\phi(x)$ of the computational scheme (\ref{eq:9}) be solved before the stopping criterion in (\ref{eq:28}) is fulfilled for the specified accuracy of the global search $\varepsilon > 0$ and the density of trials for the interval $[a,b] \subset [0,1]$ is reached
\begin{equation}\label{eq:39}
p_{ab}=\frac{n_{ab}}{b-a}
\end{equation}
where $n_{ab}$ is the number of trial points in the interval $[a,b]$ during the minimization of the function $\phi(x)$. Then, when solving the next optimization problem $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, the number of trials in the same interval $[a,b]$ instead of the maximum value
\begin{equation}\label{eq:40}
(3m/2\Delta)(b-a), m \; \text{from} \;  (\ref{eq:35}),
\end{equation}
will be limited by the value
\begin{equation}\label{eq:41}
n_{ab}' \leq (3m/2\Delta - p_{ab})(b-a)
\end{equation}
if the condition is met
\begin{equation}\label{eq:42}
\psi(x) \geq \psi(x^*)+\Delta, x \in [a,b].
\end{equation}
\end{theorem}

@@

\begin{theorem}
If the number of trials in the interval $[a,b]\subset[0,1]$ exceeds the maximum estimate (\ref{eq:41}) when solving the optimization problem $\psi(x)$ of the computational scheme (\ref{eq:9}), then when solving the next optimization problem $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, additional trials in the interval $[a,b]$ will not be performed while the condition (\ref{eq:42}) is fulfilled.
\end{theorem}

@@

It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subproblems of the same MCO problems.

>>>>

\institute{               
Victor Gergel \at
%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{gergel@unn.ru} 
\and
Evgeniy Kozinov \at
%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{evgeny.kozinov@itmm.unn.ru} 
\and
Konstantin Barkalov \at
%\at Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia \\
\email{konstantin.barkalov@itmm.unn.ru} \\
\at Mathematical Center, Lobachevsky State University of Nizhni Novgorod, Gagarin ave., 23, Nizhni Novgorod, Russia 
}

@@

For many years, global optimization has been the subject of many scientific studies. The results of research have been published in a large number of monographs -- see, for example, \cite{c17,c18,c19,c20,c21,c22,c23,c24,c25}. Making a reasonable choice of the approach for solving GO problems from the entire set of already existing global search algorithms requires taking into account the following.
\begin{itemize}
	\item Since the estimate of the global minimum is an integral characteristic of the optimization problem (the value of the minimized function at the point of the global minimum should be compared with the values of the function at all other points of the search domain), the global search methods should provide at least some degree of coverage of the search domain.
	\item Due to the ``curse of dimensionality'' reducing computational complexity requires the resulting coverage of the search domain to be heterogeneous (nonuniform), i.e. it must be dense only in the vicinity of globally optimal points.
	\item Heterogeneous coverages can only be constructed if there are any a priori assumptions about the behavior of the minimized function -- one of the most frequently used forms of such assumptions is the feasibility of the Lipschitz condition (\ref{eq:2}).
	\item Construction of heterogeneous coverages of the search domain entails the analysis of a large volume of multidimensional data (points of performed iterations of the global search and values of the minimized function at these points), which significantly increases the complexity and computational cost of the global optimization methods. Such computational complexity in the global optimization is commonly decreased by the dimensionality reduction, for example, using a diagonal extension of one-dimensional algorithms \cite{c22,c45}, the nested optimization scheme \cite{c46,c47}, the Peano space-filling curves \cite{c17,c23}.
	\item When solving multiple problems of multistage lexicographic multicriteria optimization, global search methods can reduce the complexity of calculations by using search information obtained during the calculations.
\end{itemize}

Given the requirements listed above, the proposed approach for solving MMCOlex problems is based on three key ideas:
\begin{itemize}
	\item Reduction of multidimensional multiextremal optimization to problems of one-dimensional global search by the space-filling curves;
	\item Using the global optimization methods that generate nonuniform coverages of the search domain;
	\item Intensive use of the search information obtained in the process of computations for solving multiple global optimization problems.
\end{itemize}

@@

The presence of nonlinear constraints significantly increases the complexity of global optimization problems; the resulting solutions must belong to a feasible domain
\begin{equation}\label{eq:17}
\Delta_g  = \{ x\in \Delta=[0,1]:g_i(x)\leq 0, 1 \leq i \leq m \}.
\end{equation}

@@

\begin{theorem}
If the AGCS algorithm is used to solve the problem (\ref{eq:16}), and the following conditions are satisfied:
\begin{enumerate}
	\item the functions $g_j(x)$, $1 \leq j \leq m+1$, $x\in[0,1]$, satisfy the H\"older condition with the constants $H_j$, $1 \leq j \leq m+1$,
	\item for the values $H^k_\nu$ from (\ref{eq:24}), starting from a certain step, the inequalities are valid\footnote{The values $r_{\nu} > 1$, $1 \leq \nu \leq m+1$, are the AGCS reliability parameters used for computing the interval characteristics in (\ref{eq:25}).}
\begin{equation} \label{eq:29}
r_\nu H^k_\nu > 2H_\nu, r_\nu > 1, 1 \leq \nu \leq m+1,
\end{equation}
\end{enumerate}
then the set of limit points of the sequence $\{x^k\}$ generated by the algorithm coincides with the set of global minimum points of the problem (\ref{eq:16}) with $\varepsilon=0$ in the stopping criterion (\ref{eq:28}).
\end{theorem}

@@

\begin{theorem}
Let the next minimized function $\psi(x)$ of the computational scheme (\ref{eq:9}) differ from the previous minimized function $\phi(x)$ by some limited value $\delta$, i.e.
\begin{equation}\label{eq:32}
|\psi(x)-\phi(x)| \leq \delta, x \in [0,1],
\end{equation}
and let the stopping criterion of (\ref{eq:28}) for the specified accuracy of global search $\varepsilon>0$ be fulfilled when minimizing the function $\phi(x)$. Then 
\begin{equation}\label{eq:33}
|\psi(x^*)-\phi(x_k^*)| \leq \Delta = \frac{\alpha L \varepsilon}{2} + \delta,
\end{equation}
if starting from a certain iteration $k_0>0$ the condition is fulfilled for the estimation of the Lipschitz constant $m$ and some $\alpha > 1$
\begin{equation}\label{eq:34}
m \geq L (1+ \sqrt{\frac{\alpha + 4 \beta (1 + \beta)}{\alpha - 1}}),
\end{equation}
where
\begin{equation}\label{eq:35}
m=\max{r_l H_l^k}, 1 \leq l \leq M^k, k>k_0,
\end{equation}
\begin{equation}\label{eq:36}
L=\max{L_i}, 1 \leq i \leq s, L_i\; \text{from}\; (\ref{eq:3}),
\end{equation}
\begin{equation}\label{eq:37}
\psi(x^*)=\min{\psi(x)},x \in [0,1],
\end{equation}
\begin{equation}\label{eq:38}
x_k^* = arg \min{\phi(x_k)}, 1 \leq i \leq k.
\end{equation}
\end{theorem}
The points $x_k$, $1 \leq i \leq k$, in (\ref{eq:38}) are taken from search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, and $H_l^k$, $M^k$, $r_l$ are the values of (\ref{eq:22}), (\ref{eq:24}) and (\ref{eq:29}), respectively.

@@

\begin{theorem}
Let the function $\phi(x)$ of the computational scheme (\ref{eq:9}) be minimized before the stopping criterion in (\ref{eq:28}) is fulfilled for the specified accuracy of the global search $\varepsilon > 0$ and the density of trials $p_{ab}$ for the interval $[a,b] \subset [0,1]$ is reached
\begin{equation}\label{eq:39}
p_{ab}=\frac{n_{ab}}{b-a},
\end{equation}
where $n_{ab}$ is the number of trial points in the interval $[a,b]$ during the minimization of the function $\phi(x)$. Then, when minimizing the next function $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, the number of trials in the same interval $[a,b]$ instead of the maximum value
\begin{equation}\label{eq:40}
(3m/2\Delta)(b-a), \text{where} \; m \; \text{is from} \;  (\ref{eq:35}),
\end{equation}
will be limited by the value
\begin{equation}\label{eq:41}
n_{ab}' \leq (3m/2\Delta - p_{ab})(b-a)
\end{equation}
if the condition is met
\begin{equation}\label{eq:42}
\psi(x) \geq \psi(x^*)+\Delta, x \in [a,b].
\end{equation}
\end{theorem}

@@

\begin{theorem}
If the number of trials in the interval $[a,b]\subset[0,1]$ exceeds the maximum estimate (\ref{eq:41}) when minimizing the function $\phi(x)$ of the computational scheme (\ref{eq:9}), then when minimizing the next function $\psi(x)$ using the search information $A_k$ from (\ref{eq:21}) obtained by minimizing the function $\phi(x)$, additional trials in the interval $[a,b]$ will not be performed while the condition (\ref{eq:42}) is fulfilled.
\end{theorem}

@@

It should be noted that the number of iterations of the global search is also reduced in the course of sequential solution of subproblems of the same MCO problem.
